---
title: "Linear Regression"
author: "Tapas Mishra"
date: "13/06/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



In this documents , we will look , various questions and understand , how we can solve them by regression. In process , We will understand, some R outputs 

### Question 1

You may have been told that juice drinks are “nothing but sugar.” Read in the dataset “cordial-simple.csv”,
which has the nutrition information for several different cordial brands and flavours. Consider predicting
Energy in kJ with grams of Total Sugar

a) Suppose these drinks are ideed “nothing but sugar”, and sugar contains 16.7 kJ per gram. Formulate
an appropriate null hypothesis for $$\beta_0$$ and 	$\beta_1$ that reflects this status quo. Make a scatter plot of the data; add both the line of best fit and your hypothesized line.

[latex]\hat{y} = \beta_1 x_i + \beta_0[/latex] 

### Solution

So the null hypothesis $\H_0$ for Beta_0 is $\beta_0$ = 0, and corresponding alternate hypothesis $\H_a$ for $\beta_0$ is $\beta_0$ != 0 and null hypothesis(H0)  for $\beta_1$ is Beta_1 = 16.7 and corresponding alternate hypothesis(Ha)  for $\beta_1$ is $\beta_1$ != 16.7 Now creating the plot 

```{r fig.height=5, fig.width=10}
data.cordial = read.csv("E:\\Projects\\Statistical Computation\\Statistical-Computation\\cordial-simple.csv")

fit <- lm( Energy ~ TotalSugar, data = data.cordial)

plot(Energy ~TotalSugar  , data = data.cordial,
     main = "Total Sugar Vs Energy plot", xlab = "Total Sugar in gms", ylab= "Enerygy in KJ", pch = 19)
lines(x = data.cordial$TotalSugar, y =16.7* data.cordial$TotalSugar ,col="red", lty=2)
#abline(a=0, b=16.7*data.cordial$TotalSugar, col="red", lty=2)
abline(a=coef(fit)[1], b=coef(fit)[2], col = "blue")
legend("topleft",col=c("red", "blue"), legend= c("Hypothized Line", "Best fit line"),
       lty=2:1, cex=0.8,bg='lightyellow')
```

### (b) 
Fit a linear model and report the output of the “summary” function.

###Sol b


```{r pressure, echo=FALSE}
model.Energy = lm( Energy ~ TotalSugar, data = data.cordial)
summary(model.Energy)
```

So the output of the summary shows the regression fit line has intercept of 4.5 and slope of 16.19 for Total sugar. The null hypotheis that slope of Total sugar is 0 is rejected, as we see small p-value for t-test. Therefore , this model suggests that Energy in KJ can be predicted by below equation 

Energy(in KJ) = 4.5 + 16.19*TotalSugar(in gms)


Use the output to test the hypotheses in (1) against two sided alternatives. State your conclusions.



Comment on how close the relationship is between Total Sugar and energy, citing portions of your
output in (2).


Produce and include diagnostic plots. What assumptions are they checking? Briefly comment on your
conclusions.

Are there any drinks where the kJ and Total Sugar amounts given are inconsistent, allowing that there
may be other sources of kJ?



